---
title: "Validity OF STRONGkids AND MUAC in predicting acute malnutrition in hospitalised paediatric in-patients in Accra, Ghana"
author: "Eric Komla Anku, Harriet Adu-Amoah, Joana Ainuson-Quampah"
date: "6/8/2021"
output: 
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
    latex_engine: xelatex
---



* * *


## Data and project setup


The data was collected as a purposive sample of children admitted within 24 hours of admission at both Korle Bu and Princess Marie Louis Hospital.


```{r setup, include=FALSE}
# Project setup

knitr::opts_chunk$set(echo = FALSE) # mute codes
```



```{r package-setup}
knitr::read_chunk("packages.R")
```



```{r packages}

```


```{r import-data}
nrs <- import (here("data", "nrs.csv")) # Import dataset
```



```{r data-wrangling, message=FALSE, warning=FALSE}
knitr::read_chunk("data_processing.R") # Data transformations
```


```{r transformations}

```






* * *


## Research question

Is STRONGkids and MUAC useful in predicting acute malnutrition in hospitalised pediatric patients aged 6 months and 60 months compared to weight for height z score of <-2 SD?



* * *



## Exploratory data analysis



```{r corplot,message=FALSE,cache=TRUE}
## Correlation between numerical variables
a <- nrs%>%select (age,los,weight,stature,muac,strong_score,wfh_who,muacz_who)
ggpairs(a)
```


```{r skim-dataset}
skim(nrs)
```


***



## Summary tables


### Descriptive statistics (all)



```{r summary-table-all}
#Summary statistics 


nrs%>%
  select(-c(id, diagnosis, diagnosis_cat))%>%
    tbl_summary(statistic = c(stature, 
                            muac,
                            wfh_who,
                            muacz_who
                            )
                                  ~ "{mean} ({sd})")

```



### Descriptive statistics by sex




```{r summary-table-by-sex}
nrs%>%
  select(-c(id, diagnosis, diagnosis_cat))%>%
    tbl_summary(by=sex,
                statistic = c(stature, 
                            muac,
                            wfh_who,
                            muacz_who
                            )
                                  ~ "{mean} ({sd})")
```




* * *


## Figures 

```{r figure-one}


# Plot MUAC z-score and sex

ggbarstats(
  data = nrs,
  x = sex,
  y= muac_z_cat,
  type = "parametric",
  title = "MUAC z-score category stratified by sex",
  xlab = "MUAC categories",
  legend.title = "Sex",
  package = "awtools",
  palette = "spalette"
)

```


```{r figure-two}
# Plot WHO z-score and sex

ggbarstats(
  data = nrs,
  x = sex,
  y= wfh_who_cat,
  type = "parametric",
  title = "Acute malnutrition (WHO) stratified by sex",
  xlab = "Acute malnutrition",
  legend.title = "Sex",
   package = "awtools",
  palette = "mpalette"
)

```






* * *

## Diagnostic statistics



STRONG WHO


The following is cross table for STRONG categories against nutrition status defined by WHO z-score


```{r strong-who}
table (nrs$strong.crosstab,nrs$wfh_who_cat) #STRONGKIDS WHO
```




MUAC WHO


The following is cross table for MUAC categories against nutrition status defined by WHO z-score


```{r muac-who}
table (nrs$muac.crosstab,nrs$wfh_who_cat) # MUAC WHO
```



***


*Calculating confidence intervals for diagnostic values*


STRONG SN WHO

Sensitivity of STRONGkids



```{r sensitivity-strong}
## WHO-STRONGkids

## Sensitivity
BinomCI(20, 29, conf.level = 0.95, method =  "wald")

```



STRONG SP WHO


Specificity of STRONGkids



```{r specificity-strong}
## Specificity


BinomCI(29, 67, conf.level = 0.95, method =  "wald")
```



STRONG PPV WHO


Positive predictive value of STRONGkids



```{r ppv-strong}

## Positive predictive value

BinomCI(20, 58, conf.level = 0.95, method =   "wald")
```




STRONG NPV WHO

Negative predictive value of STRONGkids


```{r npv-strong}
## Negative predictive value

BinomCI(29, 38, conf.level = 0.95, method =   "wald")
```





***



MUAC SN WHO


Sensitivity of MUAC



```{r sensitivity-muac}
## WHO-MUAC

## Sensitivity
BinomCI(13, 29, conf.level = 0.95, method =  "wald")
```




MUAC SP WHO


Specificity of MUAC


```{r specificity-muac}

## Specificity
BinomCI(62, 67, conf.level = 0.95, method =  "wald")
```



MUAC PPV WHO


Positive predictive value of MUAC



```{r ppv-muac}

## Positive predictive value

BinomCI(13, 18, conf.level = 0.95, method =   "wald")
```




MUAC NPV WHO


Negative predictive value of MUAC



```{r npv-muac}

## Negative predictive value

BinomCI(62, 78, conf.level = 0.95, method =   "wald")
```



***


## Inter-rater agreement using Cohen's Kappa


```{r cohen-kappa}
k <- import (here("data","inter_rater_dataset.csv"))

agree(k[,13:14], tolerance = 0)

kappa2(k[,13:14])
```



***


## Inferential statistics




### Weight and sex



```{r inference-weight, cache=TRUE}
### Inference

## Median weight

inference (y = weight, x = sex, data = nrs, type ="ht",
           statistic = "median",
           method = "simulation", 
           alternative = "twosided"
           )
```



### Age and sex



```{r inference-age, cache=TRUE}
## Median age

inference (y = age, x = sex, data = nrs, type ="ht",
           statistic = "median",
           method = "simulation", 
           alternative = "twosided"
)
```


### Stature and sex



```{r inference-height, cache=TRUE}
## Mean height

inference (y = stature, x = sex, data = nrs, type ="ht",
           statistic = "mean",
           method = "theoretical", 
           alternative = "twosided"
)

```


### Length of stay and sex



```{r inference-los}
## Median length of stay

inference (y = los, x = sex, data = nrs, type ="ht",
           statistic = "median",
           method = "simulation", 
           alternative = "twosided"
)
```


### MUAC and sex



```{r inference-muac}
## Mean muac

inference (y = muac, x = sex, data = nrs, type ="ht",
           statistic = "mean",
           method = "theoretical", 
           alternative = "twosided"
)

```




### STRONGkids category and sex



```{r inference-strongkids}

fisher.test(nrs$sex, nrs$strong.category) # Fisher's test for STRONGkids category
```



### MUAC categories and sex



```{r}
fisher.test(nrs$sex, nrs$muac.category) # Fisher's test for MUAC category
```


### Malnutrition prevalence and confidence interval



```{r malnutrition-prevalence}
inference (y = wfh_who_cat, data = nrs, type ="ci",
           statistic = "proportion",
           method = "theoretical", 
           alternative = "twosided",
           success = "Acute malnutrition"
)

```





