---
title: ""
author: "Eric Komla Anku, Harriet Adu-Amoah, Joana Ainuson-Quampah"
date: "6/8/2021"
output: html_document
---



* * *


## Part 1: Data


The data was collected as a purposive sample of children admitted within 24 hours of admission at both Korle Bu and Princess Marie Louis Hospital.


```{r setup, include=FALSE}
# Project setup

knitr::opts_chunk$set(echo = FALSE) # mute codes
```



```{r package-setup}

pacman::p_load (
        tidyverse, # Data management
        rio, # Importing dataset
        here, # Working directory
        janitor, # Clean data names
        googlesheets4, # Read sheets from Google sheets
        gtsummary, # Create summary tables
        skimr, # Overview of dataset
        GGally, # Correlation matrix
        tinytex, # Creating pdf documents
        finalfit, # Creating models
        ggthemes, # Plot themes
        ggstatsplot, # Inferential plots
        knitr, # Chunk options
        DescTools, # Confidence intervals
        irr, # Inter-rater agreement
        statsr #Inferential statistics
        
        
)

```



```{r import-data}
nrs <- import (here("data", "nrs.csv")) # Import dataset
```



```{r data-wrangling}

nrs <- nrs %>%
        mutate( sex = factor(sex) %>%               
                              
                       fct_recode(           
                               "Male" = "1",      
                               "Female"   = "2") %>%
                       ff_label("Sex"),  # Recode sex data into male and female 
               
             
               nationality = 
                       factor (nationality) %>%
                       
                       fct_recode("Ghanaian" = "1",
                                  "Others"  = "2") %>%
                       
                       ff_label("Nationality"), # Recode Nationality into Ghanaian and others

               
               tribe = factor (tribe) %>%
                       
                       fct_recode("Akan" = "1",
                                  "Ewe" = "2",
                                  "Ga" = "3",
                                  "Others" = "9") %>% 
                       ff_label ("Tribe"), # Recode tribe 
               
               disease = factor (diagnosis_cat) %>% 
                       fct_recode("Respiratory" = "1",
                                  "Gastrointestinal" = "2",
                                  "Infectious" = "3",
                                  "Cardiac" = "4",
                                   "Hematologic" = "5",
                                   "Tumor" = "6",
                                   "Neurologic" = "7",
                                  "Renal" = "8",
                                  "Multiple diseases" = "9",
                                   "Others" = "10"
                               
                       ) %>% 
                       
                       ff_label("Disease categories"), # Recode diagnoses into categories
               
               
               muac.category = case_when (muac < 11.5  ~ "Severe acute malnutrition",
                                          muac > 12.5 ~ "Normal nutrition status",
                                         TRUE ~ "Moderate acute malnutrition")%>%
                       ff_label("MUAC category")%>%
                       fct_relevel (c("Normal nutrition status",
                                    "Moderate acute malnutrition",
                                    "Severe acute malnutrition")), # Recode muac values into categories
                      
        
                
                muac.crosstab = if_else (muac > 12.5, "Normal",
                                        "Malnourished") %>% 
                        fct_relevel ("Normal") %>% 
                       ff_label ("MUAC two levels"), # Recode muac into two categories for crosstab 
               
               
               strong.category = factor (strong_score) %>% 
                       fct_recode("Low risk" = "0",
                                  "Medium risk" = "1",
                                  "Medium risk" = "2",
                                  "Medium risk" = "3",
                                  "High risk" = "4",
                                  "High risk" = "5") %>%
                       
                       fct_relevel(c("Low risk",
                                     "Medium risk",
                                     "High risk")), # Recode strongkids scores into risk categories
               
               
               strong.crosstab = factor (strong_score) %>% 
                       fct_recode("No risk" = "0",
                                  "At risk" = "1",
                                  "At risk" = "2",
                                  "At risk" = "3",
                                  "At risk" = "4",
                                  "At risk" = "5") %>% 
                       fct_relevel("No risk") %>% 
                       ff_label ("STRONGkids"), # Recode strongkids categories into two levels for crosstab
               
                       
               wfh_who_cat = if_else(wfh_who < - 2,
                                     "Acute malnutrition",
                                     "Normal") %>% 
                       fct_relevel("Normal") %>% 
                       ff_label ("WHO category"), # Recode wfh z-scores into acute malnutrition and none
              
              
               muac_z_cat = if_else(muacz_who < - 2,
                                    "Acute malnutrition",
                                    "Normal") %>% 
                       fct_relevel("Normal") %>% 
                       ff_label ("MUAC z-score category"), # Recode muac z-scores into acute malnutrition and none
               
              age.group = 
                      if_else(age < 24, "< 24 months",
                              "> 24 months"), # Recode age into two groups
              
              
              age.group = factor (age.group) %>%
                      ff_label("Age categories"))   # Convert age category into a factor

```







* * *


## Part 2: Research question

Is STRONGkids and MUAC useful in predicting acute malnutrition in hospitalised pediatric patients between 6 months and 60 months compared to weight for height z score of <-2 SD?



* * *



## Part 3: Exploratory data analysis



```{r corplot,message=FALSE,cache=TRUE}
## Correlation between numerical variables
a <- nrs%>%select (age,los,weight,stature,muac,strong_score,wfh_who,muacz_who)
ggpairs(a)
```


```{r skim-dataset}
skim(nrs)
```

***



## Visualisations


```{r distributions-of-variables, eval=FALSE}
# Distributions

## Histograms for numerical values and barplots for categorical variables

###Age
ggplot (nrs, aes (age)) +
  geom_freqpoly() 

###Sex
ggplot (nrs, aes (sex)) +
  geom_bar() 

###Disease category
ggplot (nrs, aes (diagnosis_cat)) +
  geom_bar(aes(fill=diagnosis_cat)) +
  coord_polar() +
  guides (fill="none")

### Length of stay
ggplot (nrs, aes (los)) +
  geom_freqpoly() 
```



```{r age-visualisations, eval=FALSE}
#What is the relationship between age and the other variables

## Age and sex
ggplot(nrs, aes(sex, age, fill = sex))+
  geom_boxplot() + 
  labs (title = "Relationship betwenn age and sex", x= "Sex of child", y= "Age of children")

## Age and diagnosis category

ggplot(nrs, aes(diagnosis_cat, age))+
  geom_boxplot() + 
  labs (title = "Relationship betwenn age and diagnosis category", x= "Disease category", y= "Age of children")

## Age and length of stay stratified by sex

ggplot(nrs, aes(age, los, colour = sex))+
  geom_jitter() + 
  labs (title = "Relationship betwenn age and lenght of stay", x= "Age of child", y= "Length of stay")

## Age and weight stratified by sex


ggplot(nrs, aes(age, weight, colour = sex))+
  geom_point() + 
  labs (title = "Relationship betwenn age and weight", x= "Age of child", y= "Weight")

## Age and stature

ggplot(nrs, aes(age, stature, colour = sex))+
  geom_point() + 
  labs (title = "Relationship betwenn age and stature", x= "Age of child", y= "Stature")

## Age and muac

ggplot(nrs, aes(age, muac, colour = sex))+
  geom_point() + 
  labs (title = "Relationship betwenn age and muac", x= "Age of child", y= "Mid upper arm circumference")

## Age and two by two muac values

ggplot(nrs, aes(muac.crosstab, age))+
  geom_boxplot(outlier.color = "red") + 
  labs (title = "Relationship betwenn age and two by two muac values", x= "MUAC categories", y= "Age of child")

## Age and Strongkids category

ggplot(nrs, aes(strong.crosstab, age))+
  geom_boxplot(outlier.color = "red") + 
  labs (title = "Relationship betwenn age and two by two STRONG values", x= "Strong categories", y= "Age of child")


## Age and Weight for height WHO

ggplot(nrs, aes(age, wfh_who, colour = sex))+
  geom_point() + 
  labs (title = "Relationship betwenn age and weight for height (WHO)", x= "Age of child", y= "Weight for height")

## Age and MUAC z score

ggplot(nrs, aes(age, muacz_who, colour = sex))+
  geom_point() + 
  labs (title = "Relationship betwenn age and muac z scores", x= "Age of child", y= "Mid upper arm circumference z scores")


## Age and WHO wfh categories

ggplot(nrs, aes(wfh_who_cat, age))+
  geom_boxplot(outlier.color = "red") + 
  labs (title = "Relationship betwenn age and two by WHO malnourished categories", x= "Acute malnutrition by WHO", y= "Age of child")

## Age and MUAC z score categories

ggplot(nrs, aes(muac_z_cat, age))+
  geom_boxplot(outlier.color = "red") + 
  labs (title = "Relationship betwenn age and MUAC z score categories", x= "Acute malnutrition by MUAC", y= "Age of child")



```




```{r summary-table}
#Summary statistics 


nrs%>%
  select(-c(id, diagnosis, diagnosis_cat))%>%
    tbl_summary()

```









```{r eval=FALSE}
nrs%>%filter(muac_z_cat=="Acute malnutrition")%>%
  group_by(age.group)%>%
  count()
```




Descriptive information appendix



```{r eval=FALSE}
nrs %>% group_by(sex)%>%
  summarise(median_age = median (age),
            IQR_age = IQR (age),
            median_los = median (los),
            IQR_los = IQR (los),
            median_wt = median (weight),
            IQR_wt = IQR (weight),
            mean_ht = mean (stature),
            sd_ht = sd (stature),
            mean_muac = mean (muac),
            sd_muac = sd (muac))

nrs %>% group_by(sex,wfh_cdc_cat,
                )%>%
  summarise (n = n())


nrs %>% group_by(sex,wfh_who_cat,
                )%>%
  summarise (n = n())
 

nrs %>% group_by(sex,muac_z_cat,
                )%>%
  summarise (n = n())
 
 nrs %>% group_by(sex,muac_cat_r,
                )%>%
  summarise (n = n())
 
 
  nrs %>% group_by(sex,strong_score_cat,
                )%>%
  summarise (n = n())                

```


* * *


```{r}
ggplot (nrs, aes (strong.category, fill = sex)) +
  geom_bar (position = "dodge") + labs (title = "STRONGkids risk category by sex", x = "Risk category", y = "Frequency") + scale_fill_discrete(name= "Sex")


# plot strongkids
ggbarstats(
  data = nrs,
  x = sex,
  y = strong.crosstab,
  type = "parametric",
  title = "STRONGkids category stratified by sex",
  xlab = "STRONGkids categories",
  legend.title = "Sex",
  package = "basetheme",
  palette = "ink"
)


# plot muac

ggbarstats(
  data = nrs,
  x = sex,
  y = muac.crosstab,
  type = "parametric",
  title = "MUAC category stratified by sex",
  xlab = "MUAC categories",
  legend.title = "Sex",
  package = "beyonce",
  palette = "X1"
)




# Plot MUAC categories
ggbarstats(
  data = nrs,
  x = sex,
  y= muac_z_cat,
  type = "parametric",
  title = "MUAC z-score category stratified by sex",
  xlab = "MUAC categories",
  legend.title = "Sex",
  package = "awtools",
  palette = "spalette"
)


# Plot WHO
ggbarstats(
  data = nrs,
  x = sex,
  y= wfh_who_cat,
  type = "parametric",
  title = "Acute malnutrition (WHO) stratified by sex",
  xlab = "Acute malnutrition",
  legend.title = "Sex",
   package = "awtools",
  palette = "mpalette"
)

```



MUAC z score vrs WHO z score
```{r}
## MUAC z-score vrs WHO z-score

ggscatterstats(
  data = nrs,
  x = muacz_who,
  y= wfh_who,
  type = "parametric",
  title = "Acute malnutrition WHO stratified by sex",
  xlab = "MUAC z-scores",
  ylab = "WHO z-scores"
)

```


* * *

## Part 6: Conclusion



STRONG WHO


```{r strongWHO}
table (nrs$strong.crosstab,nrs$wfh_who_cat) #STRONGKIDS WHO
```




MUAC WHO


```{r}
table (nrs$muac.crosstab,nrs$wfh_who_cat) # MUAC WHO
```



***


*Calculating confidence intervals for diagnostic values*


STRONG SN WHO

```{r}
## WHO-STRONGkids

## Sensitivity
BinomCI(20, 29, conf.level = 0.95, method =  "wald")

```



STRONG SP WHO


```{r}
## Specificity
BinomCI(29, 67, conf.level = 0.95, method =  "wald")
```



STRONG PPV WHO



```{r}

## Positive predictive value

BinomCI(20, 58, conf.level = 0.95, method =   "wald")
```


STRONG NPV WHO


```{r}
## Negative predictive value

BinomCI(29, 38, conf.level = 0.95, method =   "wald")
```





MUAC SN WHO


```{r}
## WHO-MUAC

## Sensitivity
BinomCI(13, 29, conf.level = 0.95, method =  "wald")
```




MUAC SP WHO


```{r}
## Specificity
BinomCI(62, 67, conf.level = 0.95, method =  "wald")
```



MUAC PPV WHO


```{r}
## Positive predictive value

BinomCI(13, 18, conf.level = 0.95, method =   "wald")
```




MUAC NPV WHO


```{r who-muac}

## Negative predictive value

BinomCI(62, 78, conf.level = 0.95, method =   "wald")
```



***


*Inter-rater agreement using Cohen's Kappa*


```{r cohen-kappa}
k <- import (here("data","inter_rater_dataset.csv"))

agree(k[,13:14], tolerance = 0)

kappa2(k[,13:14])
```



***


*INFERENTIAL STATISTICS*



```{r inference, cache=TRUE}
### Inference

## Median weight

inference (y = weight, x = sex, data = nrs, type ="ht",
           statistic = "median",
           method = "simulation", 
           alternative = "twosided"
           )

## Median age

inference (y = age, x = sex, data = nrs, type ="ht",
           statistic = "median",
           method = "simulation", 
           alternative = "twosided"
)

## Mean height

inference (y = stature, x = sex, data = nrs, type ="ht",
           statistic = "mean",
           method = "theoretical", 
           alternative = "twosided"
)


## Median length of stay

inference (y = los, x = sex, data = nrs, type ="ht",
           statistic = "median",
           method = "simulation", 
           alternative = "twosided"
)

## Mean muac

inference (y = muac, x = sex, data = nrs, type ="ht",
           statistic = "mean",
           method = "theoretical", 
           alternative = "twosided"
)




## muac_z score

inference (y = muac_z_cat, x = sex, data = nrs, type ="ht",
           statistic = "proportion",
           method = "theoretical", 
           alternative = "twosided",
           success = "Acute malnutrition"
)

## WHO WFH


inference (y = wfh_who_cat, x = sex, data = nrs, type ="ht",
           statistic = "proportion",
           method = "theoretical", 
           alternative = "twosided",
           success = "Acute malnutrition"
)





inference (y = wfh_who_cat, data = nrs, type ="ci",
           statistic = "proportion",
           method = "theoretical", 
           alternative = "twosided",
           success = "Acute malnutrition"
)


```


